---
title: "MultinomialClasification"
author: "Julian Avila-Jimenez"
date: "3/8/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE,echo = TRUE, warning = FALSE,
                      message = FALSE, dpi= 180, fig.width = 8,
                      fig.height = 5)
library(tidyverse)
```

## Get the data from [Volcano](https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-05-12) data from TidyTuesday 2020-05-12.
```{r}
Volcano<- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/volcano.csv")

Volcano
```

## Explore the Volcano dataset.
```{r}
Volcano %>% 
  count(primary_volcano_type, 
        sort = T) %>% 
  knitr::kable()
```

Here we see the different types of volcano classes, maybe we can do a MultiClass Classification (multinomial) to determine the type of volcano.

26 classes is too much for under 1000 registries, so we can reduce the classes to a few less to work.


```{r}
Volcano_df<- Volcano %>% 
  transmute(volcano_type= case_when(str_detect(primary_volcano_type,
                                               "Stratovolcano") ~ "Stratovolcano  ",
                                    str_detect(primary_volcano_type, "Shield") ~ "Shield",
                                    str_detect(primary_volcano_type, "Volcanic field") ~ "Volcanic field", 
                                    TRUE ~"Other"),
            volcano_number, latitude, longitude, 
            elevation, tectonic_settings, major_rock_1) %>% 
  mutate_if(is.character, factor)
```

Let's see the spatial distribution of the volcanoes 
```{r}
world<- map_data("world")

ggplot()+
  geom_map(data= world, map= world,
           aes(long, lat, map_id = region), 
               color= "white", fill= "gray50", alpha= 0.2)+
  geom_point(data= Volcano_df,
             aes(longitude, latitude, color= volcano_type), 
             alpha= 0.8)+ 
  scale_color_brewer(palette="Dark2")+
  labs(title= "Volcanoes geographic distribution\nby Volcano type")+
  theme_bw()+
  theme(legend.title = element_blank())
```

## Modeling Proces.
There is no sufficient data to split into train and test, so we can do a bootstrap

```{r}
library(tidymodels)

Volcano_boot<- bootstraps(Volcano_df)
Volcano_boot
```

```{r}
Volcano_df %>% 
  count(volcano_type)
```

The number of each type of volcanoes is too unbalanced so we have to deal with it with a SMOTE algorithm of the [themis](https://github.com/tidymodels/themis) library.

To do this first we have create a Dummy variable to transform the factors to numeric and center scale the numeric variables. 
```{r}
Volcano_df %>% 
  count(tectonic_settings) #too much categories!

Volcano_df %>% 
  count(tectonic_settings) #Same, too much categories!
```

### Preprocesing.
Let's make the recipe!
```{r}
library(themis)

Volcano_rec<-recipe(volcano_type ~ ., data= Volcano_df ) %>% #Predict Volcano type with everything else
  update_role(volcano_number, new_role = "Id") %>% 
  step_other(tectonic_settings) %>%  #collapses the most infrequent levels to "Other"
  step_other(major_rock_1) %>% #same as before
  step_dummy(tectonic_settings, major_rock_1) %>%  #creates Dummy variables
  step_zv(all_predictors()) %>% #remove variables with zero variance
  step_normalize(all_predictors()) %>% #centerscale all predictors
  step_smote(volcano_type) #generate new examples of the minority class using nearest neighbors of these cases
  
Volcano_prep<- prep(Volcano_rec) #train the recipe
Volcano_prep
```
Notice how many new info we have, the new rows are because the oversampling and columns by the new Dummy variables
```{r}

juice(Volcano_prep)
juice(Volcano_prep) %>% 
  count(volcano_type) %>% #same number of observations by the oversampling of step_ smote()
  knitr::kable()
```

### Build a model.
First let's try a Random forest

Advantages:
- Is versatile, can be used for prediction and classification problems 
- No need to fine tuning parameters with a significant number of trees

we are going to fit a model using the advantages of [TidyModels](https://www.tidymodels.org/), using the [ranger](https://rdrr.io/cran/ranger/man/ranger.html) engine, a fastest implementation of Random Forest algorithm, and setting a workflow 

```{r}
rf_spec<-
  rand_forest(trees=1000) %>%
  set_mode("classification") %>% 
  set_engine("ranger")

Volcano_wf<- 
  workflow() %>% 
  add_recipe(Volcano_rec) %>% 
  add_model(rf_spec)

Volcano_wf
```

Results
```{r}
Volcano_res<-
  fit_resamples(
  Volcano_wf,
  resamples= Volcano_boot, #we have 25 resamples 
  control= control_resamples(save_pred = TRUE)
)

Volcano_res
```

## Explore the results
Look that the estimators are adjusted to a classification problem 
```{r}
Volcano_res %>% 
  collect_metrics() %>% 
  knitr::kable()

```
The classification don't perform very well, just above the random probabilities but maybe influenced by the small dataset...

```{r}
Volcano_res %>% 
  collect_predictions() %>% 
  conf_mat(volcano_type, .pred_class)
```

Positive predictive value [Wikipedia](https://en.wikipedia.org/wiki/Positive_and_negative_predictive_values)
```{r}
Volcano_res %>% 
  collect_predictions() %>% 
  ppv(volcano_type, .pred_class) %>% 
  knitr::kable()

```

### Now let's explore the variable importance 

```{r}
library(vip)
rf_spec %>% 
  set_engine("ranger", 
             importance= "permutation") %>% 
  fit(
    volcano_type ~ ., 
    data= juice(Volcano_prep) %>%
    select(-volcano_number) %>% 
    janitor:: clean_names() 
  ) %>% 
  vip(geom="col",fill="cyan4")

```
it's clear that the longitude and latitude are the most important variables of the classification process 


## Now a final map

```{r}
Volcano_predictions<- 
  Volcano_res %>% 
  collect_predictions() %>% 
  mutate(Correct = volcano_type == .pred_class) %>% 
  left_join(Volcano_df %>% 
              mutate(.row= row_number()))
Volcano_predictions
```

```{r}
ggplot()+
  geom_map(data=world,map=world,
           aes(long, lat, map_id= region),
           color="white", fill="gray50", alpha=0.2)+
  stat_summary_hex(data= Volcano_predictions, 
                   aes(longitude, latitude, z= as.integer(Correct)),
                   fun= "mean", alpha= 0.7, bins= 60)+
  scale_fill_gradient(high= "cyan3", labels= scales::percent)+
  theme_bw()+
  labs(title= "Volcanoes geographic distribution",
       subtitle= "classification performance",
       fill=" Percent classified \ncorrectly")
```

